\addchap{Preface}
\begin{refsection}

This handbook is the result of discussions which began in late 2018 among many of the handbook authors, who felt that a wide-coverage handbook of LFG would be very useful for practitioners of LFG as well as for people wanting to learn more about various aspects of the theory.  We originally planned to finish the handbook by mid-2021, but, as with many things, the COVID-19 pandemic got in the way, and the number of originally planned chapters had to be reduced. We are nevertheless delighted with the results, and we anticipate that the handbook will be an essential resource for work in LFG for many years to come.

This handbook has been a community effort in the best sense, with involvement and input from handbook contributors at every stage of the process. I have been extremely grateful to all of the contributors for their flexibility, good humor, perseverance, and very hard work.   At an early stage, Annie Zaenen and Nigel Vincent provided very helpful suggestions for the overall organization of the handbook.  As always, Tracy Holloway King provided sage advice, and help with Latex conversion at crucial moments. Bozhil Hristov went far beyond the call of duty in proofreading many of the handbook chapters.  We have also benefited greatly from the helpful advice and expert technical assistance of the editorial team, Sebastian Nordhoff and Felix Kopecky, and the vision of Stefan M\"uller and his team in establishing Language Science Press as an essential resource for linguistics.  
 
Reviewers' suggestions and advice were crucial in improving the content and presentation of the chapters. All of the handbook authors also acted as reviewers for other handbook chapters, and we are grateful to them for taking on this work.  Most authors reviewed two other handbook chapters, but the following authors took on reviewing responsibility for more than two chapters, and we are grateful to them for this extra effort: Ash Asudeh, Avery Andrews, Dag Haug, John Lowe, Peter Sells, and Tracy Holloway King.  We also relied on a large team of external reviewers who provided very useful comments and feedback.  Most external reviewers reviewed only one chapter, but we extend special thanks to those who reviewed more than one chapter: Amanda Thomas, Doug Arnold, and especially Joey Lovestrand, who contributed three reviews.  We are particularly grateful to a team of Oxford graduate students who provided very detailed and helpful comments on the introductory chapters:  Shuting Chen, Wilson Lui, Yiwei Si, Eden Watkins, and Xiulin Yang.  For very helpful comments on the glossary, we are grateful to Ron Kaplan, Elaine Ui Dhonnchadha, Bozhil Hristov, Rachel Nordlinger, Hannah Booth, Nigel Vincent, Gy\"orgy R\'akosi, Alex Alsina, and Tibor Laczk\'o. We are also grateful to the Press's very competent and helpful team of proofreaders for their careful work on all of the chapters. 

As always, I am grateful to Ken Kahn for all kinds of support as this handbook came together.\\\bigskip

Oxford, October 2023\hfill\hbox{Mary Dalrymple}

\end{refsection}

